KPIs and Visualisations

### 2022-07-31 Report/KPIs and Visualisations.Rmd
Updating.

#### Årsresultat

@@

#### Balance

@@

```{r}
# Reactive function to check for currency and use currency code or ?
currency_scale <- reactive({
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )
})
```

@@

```{r}
# Reactive function to create plotly caption (ggplot captions and subtitles
# are lost in the conversion) (shamelessly stolen from the Internet, with
# minor modifications).
# Checks for uniform currency code and prints a warning
# if there are multiple currency codes.
plotly_caption <- reactive({
  list(
    x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
      "", "Note: not all values are denominated in the same currency"
    ),
    showarrow = F, xref = "paper", yref = "paper",
    xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
    font = list(size = 11)
  )
})
```

@@

```{r}
# Reactive function to create the maximum value for revenue and expenses plots
maximum <- reactive({
  # reading the Excel sheet anew
  read_excel(input$excel$datapath) |>
    # removing the firm and sub-category columns
    select(-Firm, -`Sub-category`) |>
    # filtering to retain only revenue and expenses entries
    filter(Category == "Revenue" | Category == "Expenses") |>
    # grouping and summarising to calculate maximum value
    group_by(Year, Category) |>
    summarise(
      sum = sum(Value, na.rm = TRUE)
    ) |>
    # Creating list of sums
    pull() |>
    # selecting the highest value
    max()
})
```

@@

```{r data}
## Function to read html file --------------------------------------------------
read_proff <- function(file, start, end) {
  # Specifying encoding to deal with occasional errors
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    # removing empty column, using all_of(argument)
    select( 
      REGNSKAPSPERIODE:all_of(start)
    ) |> 
    # removing white space
    mutate( 
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter( 
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter( 
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer( 
      all_of(end):all_of(start),
      names_to = "year"
    ) |> 
    # creating a currency code column and turning current currency codes into
    # NAs
    mutate( 
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    # filling the currency column
    fill(
      valutakode,
      .direction = "updown"
    ) |>
    # removing dates and redundant currency
    filter( 
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate( 
      # removing white space in numbers
      value = str_replace_all(value, "\\s", ""),
      # turning numbers into true numbers
      value = as.numeric(value),
      # removing values that are years
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    # removing any remaining duplicate rows
    distinct() |> 
    # pivoting to create 'tidy' variables
    pivot_wider( 
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    # tidying names
    clean_names() |> 
    # removing headings from the table
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |> 
    # making it clear what lonn refers to
    rename( 
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # arranging by year for practical display
    arrange(desc(year)) 
}

# Creating data frame
htmldata <- read_proff("Data/Kreftforeningen.html", "2002", "2021")

## Function to read Excel file -------------------------------------------------
widen_excel <- function(file) {
  read_excel(file) |>
    # uniting Category and Name to deal with multiple 'Other'
    unite(
      entry, c(Category, Name),
      sep = "_"
    ) |>
    # deleting Sub-category and Firm
    select(-`Sub-category`, -Firm) |> 
    # pivoting to create 'tidy' variables
    pivot_wider(names_from = entry, values_from = Value) |>
    # cleaning names
    clean_names() 
}

# Creating data frame
exeldata <- widen_excel("Data/OPXfinans_final.xlsx")
```

>>>>

#### Net income/loss

@@

#### Total equity/liabilities

@@

```{r}
# Reactive function to check for currency and use currency code or ?
currency_scale <- reactive({
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(html_currency()$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(unique(html_currency()$valutakode), " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )
})
```

@@

```{r}
# Reactive function to create plotly caption (ggplot captions and subtitles
# are lost in the conversion) (shamelessly stolen from the Internet, with
# minor modifications).
# Checks for uniform currency code and prints a warning
# if there are multiple currency codes.
plotly_caption <- reactive({
  list(
    x = 1, y = -0.1,
    text = ifelse(length(unique(html_currency()$valutakode)) == 1,
      "", "Note: not all values are denominated in the same currency"
    ),
    showarrow = F, xref = "paper", yref = "paper",
    xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
    font = list(size = 11)
  )
})
```

@@

```{r}
# Reactive function to create the maximum value for revenue and expenses plots
# the tests are probably not necessary, but included to be on the safe side
maximum <- reactive({
  if (datatest() == TRUE || setequal(
    names(exceldata()),
    c("year", "firm", "category", "sub_category", "name", "value")
  )) {
    NA
  } else {
    excel_test() |>
      # removing the firm and sub-category columns
      select(-firm, -sub_category) |>
      # filtering to retain only revenue and expenses entries
      filter(category == "Revenue" | category == "Expenses") |>
      # grouping and summarising to calculate maximum value
      group_by(year, category) |>
      summarise(
        sum = sum(value, na.rm = TRUE)
      ) |>
      # Creating list of sums
      pull() |>
      # selecting the highest value
      max()
  }
})
```

@@

```{r data}
# Data reading functions =======================================================

## Functions to read html file -------------------------------------------------
html_test_create <- function(file, start, end) {
  # Reading file and node
  htmltest <- read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame")

  ### A series of tests to catch problems before they spit out errors-----------
  # Checking if node exists
  if (class(htmltest) == "xml_node") {
    htmltable <- htmltest |>
      html_table()
  } else {
    validate("Please upload the specified type of html document.")
  }

  # Basic check if the table is correct
  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
    htmlready <- htmltable
  } else {
    validate("Please upload the specified type of html document.")
  }

  # Testing if the correct years have been entered
  if (as.character(all_of(start)) %in% names(htmlready) &&
    as.character(all_of(end)) %in% names(htmlready)) {
    # Creating the first part of the data frame
    read_proff_one(htmltable, start, end)
  } else {
    validate("Please ensure you have entered the correct start and end years.")
  }
}

### First html transformation function, invoked by html_test_create() ----------
# (stopping to allow checks for differing currency codes)
read_proff_one <- function(object, start, end) {
  object |>
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    # creating a currency code column and turning current currency codes into
    # NAs
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    # filling the currency column to avoid having to deal with NAs later
    fill(
      valutakode,
      .direction = "downup"
    )
}

### Second html transformation function, creating the data frame ---------------
read_proff_two <- function(object) {
  object |>
    # removing superfluous currency column
    select(-valutakode) |>
    # removing dates and redundant currency
    filter(
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate(
      # removing white space in numbers
      value = str_replace_all(value, "\\s", ""),
      # turning numbers into true numbers
      value = as.numeric(value),
      # removing values that are years
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    # removing any remaining duplicate rows
    distinct() |>
    # pivoting to create 'tidy' variables
    pivot_wider(
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    # tidying names
    clean_names() |>
    # removing headings from the table
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |>
    # making it clear what lonn refers to
    rename(
      lederlonn = lonn
    ) |>
    # Creating new variables for plots (red/green profit/loss)
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # arranging by year for practical display
    arrange(desc(year))
}

## Functions to read Excel file ------------------------------------------------
### First Excel transformation -------------------------------------------------
# creating df used by NGO title function and maximum scale function
widen_excel_test <- function(file) {
  # Reading Excel file
  read_excel(file) |>
    # cleaning names to enable later operations
    clean_names()
}

### Second Excel function - creating primary Excel data frame-------------------
widen_excel <- function(object) {
  # testing if all columns are present
  if (setequal(
    names(object),
    c("year", "firm", "category", "sub_category", "name", "value")
  )) {
    object |>
      # uniting category and name to deal with multiple 'Other'
      unite(
        entry, c(category, name)
      ) |>
      # deleting Sub-category and Firm
      select(-sub_category, -firm) |>
      # pivoting to create 'tidy' variables
      pivot_wider(names_from = entry, values_from = value) |>
      # cleaning names
      clean_names()
    # Notification if all columns are not present
  } else {
    validate("Please upload data using the specified Excel template.")
  }
}

# Data frames ================================================================
## Creating data frame from html file-----------------------------------------
html_currency <- reactive({
  # validating file
  validate(need(
    tools::file_ext(input$html$datapath) == "html",
    "Please upload an html file."
  ))

  # Testing and creating data frame
  html_test_create(
    input$html$datapath,
    as.character(input$start), as.character(input$end)
  )
})

htmldata <- reactive({
  read_proff_two(html_currency())
})

## Validating and creating long version of Excel file, if possible------------
excel_test <- reactive({
  # validating file
  validate(need(
    tools::file_ext(input$excel$datapath) == "xlsx",
    "Please upload an Excel file."
  ))

  widen_excel_test(input$excel$datapath)
})

## Creating the primary Excel data frame--------------------------------------
exceldata <- reactive({
  widen_excel(excel_test())
})
```
